---
title: "Stat 330 Final"
author: "Elizabeth Mitchell"
date: "12/11/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
The mountain pine beetle (MPB), once a helpful agent in the forest life cycle has become a harmful epidemic in the Rocky Mountains. In the last few decades there haven't been enough extremely cold winters to control the size of the beetle population. With the surplus population, beetles that in the past burrowed into weakened or dead trees are now attacking healthy trees as well and damaging entire forests. We want to examine infested areas and their features including the lowest temperature in January, highest temperature in August, the angle of the mountain, elevation and preciptation. We want to determine which factors contribute to risk of infestation as well as be able to predict regions that are at risk for future investation.

```{r}
library(ggplot2)
library(gmodels)
library(bestglm)
library(pROC)
library(stargazer)
library(dplyr)
library(car)
library(stargazer)

pine.beetles <- read.csv("PineBeetle.csv", sep = ",", head = TRUE)

pine.beetles$Infested <- as.numeric(pine.beetles$Infested)

pine.beetles$Infested[pine.beetles$Infested == 1] <- 0
pine.beetles$Infested[pine.beetles$Infested == 2] <- 1

pine.beetles <- mutate(pine.beetles, region = ifelse(NW == "Yes", "NW", ifelse(SC == "Yes", "SC", ifelse(SE == "Yes", "SE", ifelse(SW == "Yes", "SW", ifelse(WC == "Yes", "WC", ifelse(NC == "Yes", "NC", ifelse(EC == "Yes", "EC", NA))))))))

ggplot(data = pine.beetles, mapping = aes(x = January, y = Infested)) + geom_point() + geom_jitter(height = .2, width = 1) + xlab("January Min Temp")

ggplot(data = pine.beetles, mapping = aes(x = August_max, y = Infested)) + geom_point() + xlab("August Max Temp")+ geom_jitter(height = .2, width = 1)

ggplot(data = pine.beetles, mapping = aes(x = Slope, y = Infested)) + geom_point() + geom_jitter(height = .1, width = 1)

ggplot(data = pine.beetles, mapping = aes(x = Elev, y = Infested)) + geom_point() + geom_jitter(height = .2, width = 1) + xlab("Elevation")

ggplot(data = pine.beetles, mapping = aes(x = Precip, y = Infested)) + geom_point() + geom_jitter(height = .2, width = 1) + xlab("Precipitation (inches)")

addmargins(table(pine.beetles$region, pine.beetles$Infested))
```

```{r}
pine.beetles$region <- NULL

beetle_model <- bestglm(pine.beetles, IC = "AIC", method = "exhaustive", family = binomial)

beetle_model <- beetle_model$BestModel
summary(beetle_model)
```

$Y_{i}$ $\stackrel{ind}{\sim}$ Bern$(p_{i})$
log$p_{i}\over1- p_{i}$ = $\beta_{0} + \beta_{1}x_{i1} + \beta_{2}x_{i2} + \beta_{3}x_{i3} + \beta_{4}x_{i4} + \beta_{5}x_{i5} + \beta_{6}x_{i6} + \beta_{7}x_{i7} + \beta_{8}x_{i8}$

```{r}
#check assumptions

scatter.smooth(pine.beetles$January, pine.beetles$Infested, xlab = "January Min Temp", ylab = "Infested")
scatter.smooth(pine.beetles$August_max, pine.beetles$Infested, xlab = "August Max Temp", ylab = "Infested")
with(pine.beetles, lines(loess.smooth(August_max, Infested), col = "red"))

scatter.smooth(pine.beetles$Slope, pine.beetles$Infested, xlab = "Slope", ylab = "Infested")

plot(pine.beetles$Elev, pine.beetles$Infested, xlab = "Elevation", ylab = "Infested")
with(pine.beetles, lines(loess.smooth(Elev, Infested), col = "red"))
scatter.smooth(pine.beetles$Precip, pine.beetles$Infested, xlab = "Preciptation", ylab = "Infested")
```


```{r}
(exp(confint(beetle_model))- 1)*100

stargazer(beetle_model, out = "beetles.html", ci = T)

```

```{r}
pred.probs <- predict.glm(beetle_model, type="response")

thresh <- seq(from=0, to=1, length=10000) 

misclass <- rep(NA,length=length(thresh)) #Empty vector to hold misclassification rates

for(i in 1:length(thresh)) {
  #If probability greater than threshold then 1 else 0
  my.classification <- ifelse(pred.probs>thresh[i], 1, 0)
  
  # calculate the pct where my classification not eq truth
  misclass[i] <- mean(my.classification!=pine.beetles$Infested)
}

(c <- thresh[which.min(misclass)])

ggplot(mapping = aes(x = thresh, y = misclass)) + geom_point() + geom_vline(xintercept =c) + xlab("Threshold") + ylab("Missclassifications")

my.roc <- roc(pine.beetles$Infested, pred.probs)
ggplot() + geom_line(aes(x=1-my.roc[["specificities"]], y=my.roc[["sensitivities"]])) + geom_abline(intercept=0, slope=1)
(auc(my.roc))
```



```{r}
pred.class <- ifelse(pred.probs > c, 1, 0)

conf.matrix <- addmargins(table(pred.class, pine.beetles$Infested))

(sens <- conf.matrix[2,2]/conf.matrix[3, 2])
(spec <- conf.matrix[1, 1]/conf.matrix[3, 1])
(PPV <- conf.matrix[2,2]/conf.matrix[2, 3])
(NPV <- conf.matrix[1, 1]/conf.matrix[1, 3])

conf.matrix
```


```{r}
## Choose number of CV studies to run in a loop & test set size
n.cv <- 1000
n.test <- round(.1*nrow(pine.beetles))

## Set my threshold for classifying
cutoff <- c
  
## Initialize matrices to hold CV results
sens <- rep(NA, n.cv)
spec <- rep(NA, n.cv)
ppv <- rep(NA, n.cv)
npv <- rep(NA, n.cv)
auc <- rep(NA, n.cv)

## Begin for loop
for(cv in 1:n.cv){
  
  test.obs <- sample(1:nrow(pine.beetles), n.test)
  test.set <- pine.beetles[test.obs,]
  train.set <- pine.beetles[-test.obs,]
  
  ## Fit best model to training set
  train.model <- glm(Infested ~ January + August_max + Slope + Elev + Precip + NC + SE + SW, data=train.set, family=binomial)
  
  ## Use fitted model to predict test set
  pred.probs <- predict.glm(train.model, newdata=test.set, type="response")

  ## Classify according to threshold
  test.class <- ifelse(pred.probs>cutoff, 1, 0)
  
  ## Create a confusion matrix
  conf.mat <- addmargins(table(factor(test.set$Infested, levels = c(0, 1)), factor(test.class, levels = c(0, 1))))

  ## Pull of sensitivity, specificity, PPV and NPV using bracket notation
  sens[cv] <- conf.mat[2,2]/conf.mat[2,3]
  spec[cv] <- conf.mat[1,1]/conf.mat[1,3]
  ppv[cv] <- conf.mat[2,2]/conf.mat[3,2]
  npv[cv] <- conf.mat[1,1]/conf.mat[3,1]
    

}

(mean(sens))
(mean(spec))
(mean(ppv))
(mean(npv))
```

```{r}
ten.years <- data.frame(January = c(-13.98, -17.8, -17.27, -12.52, -15.99, -11.97, -15.75, -16.19, -17.87, -12.44), August_max = c(15.89, 18.07, 16.74, 18.06, 18.23, 15.81, 16.85, 16.51, 17.84, 16.96), Slope = rep(18.07, 10), Elev = rep(1901.95, 10), Precip = c(771.13, 788.54, 677.63, 522.77, 732.32, 615.96, 805.9, 714.57, 740.5, 801.22), SE = rep("Yes", 10), NC = rep("No", 10), SW = rep("No", 10))

predict.glm(beetle_model, newdata = ten.years, type = "response")

mean(predict.glm(beetle_model, newdata = ten.years, type = "response"))
```

